---
title: "Amani Streams"
format: html
editor: visual
---

```{r}

library(tidyverse)
library(ggplot2)
library(dplyr)
library(gridExtra)

```

## Plotting the Data

```{r}

data <- read.csv("AmaniStreamSummaryData.csv")

head(data)

```

```{r}

summary_data <- data |>
  mutate(StreamType = recode(StreamType, "Forest" = "Control")) |>
  group_by(StreamType) |>
  summarise(
    mean_Family_Rich = mean(Family_Rich),
    se_Family_Rich = sd(Family_Rich) / 2,  # since n = 4, sqrt(4) = 2
    mean_EPT_Spp = mean(EPT_Spp),
    se_EPT_Spp = sd(EPT_Spp) / 2,  
    mean_EstSpp = mean(EstSpp),
    se_EstSpp = sd(EstSpp) / 2, 
    mean_ObsSpp = mean(ObsSpp),
    se_ObsSpp = sd(ObsSpp) / 2
  )


head(summary_data)

```

```{r}

# Panel 1: Family Richness 
plot1 <- ggplot(summary_data, aes(x = StreamType, y = mean_Family_Rich, fill = StreamType)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5, width = 0.4) + 
  geom_errorbar(aes(ymin = mean_Family_Rich - se_Family_Rich, ymax = mean_Family_Rich + se_Family_Rich), 
                width = 0.25, position = position_dodge(0.8), color = "black") +
  labs(title = NULL, x = NULL, y = "Family Richness") +
  scale_fill_manual(values = c("darkgrey", "white")) +
  scale_y_continuous(limits = c(0, 25), breaks = seq(0, 25, by = 5), expand = c(0, 0)) +  
  scale_x_discrete(labels = c("Control" = "Control", "Tea" = "Tea")) + 
  theme_minimal(base_size = 14) + 
  theme(axis.text.x = element_text(size = 10), 
        axis.title.x = element_blank(),  
        axis.title.y = element_text(size = 10), 
        axis.text.y = element_text(size = 10), 
        plot.title = element_text(size = 12), 
        legend.position = "none",  
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", size = 0.5),
        axis.ticks.y = element_line(color = "black", size = 0.5)) + 
  coord_cartesian(clip = "off")



```

```{r}
# Panel 2: EPT Richness
plot2 <- ggplot(summary_data, aes(x = StreamType, y = mean_EPT_Spp, fill = StreamType)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5, width = 0.4) +
  geom_errorbar(aes(ymin = mean_EPT_Spp - se_EPT_Spp, ymax = mean_EPT_Spp + se_EPT_Spp), 
                width = 0.25, position = position_dodge(0.8), color = "black") +
  labs(title = NULL, x = NULL, y = "EPT Richness") +
  scale_fill_manual(values = c("darkgrey", "white")) +
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 2), expand = c(0, 0)) +  
  scale_x_discrete(labels = c("Control" = "Control", "Tea" = "Tea")) +
  theme_minimal(base_size = 14) + 
  theme(axis.text.x = element_text(size = 10),  
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10), 
        plot.title = element_text(size = 12),  
        legend.position = "none",  
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", size = 0.5),
        axis.ticks.y = element_line(color = "black", size = 0.5)) + 
  coord_cartesian(clip = "off")


```

```{r}

# Panel 3: Estimated Number of Species 
plot3 <- ggplot(summary_data, aes(x = StreamType, y = mean_EstSpp, fill = StreamType)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5, width = 0.4) +
  geom_errorbar(aes(ymin = mean_EstSpp - se_EstSpp, ymax = mean_EstSpp + se_EstSpp), 
                width = 0.25, position = position_dodge(0.8), color = "black") +
  labs(title = NULL, x = NULL, y = "Estimated Number of Species") +
  scale_fill_manual(values = c("darkgrey", "white")) +
  scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10), expand = c(0, 0)) + 
  scale_x_discrete(labels = c("Control" = "Control", "Tea" = "Tea")) +
  theme_minimal(base_size = 14) + 
  theme(axis.text.x = element_text(size = 10),  
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 10), 
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size = 12),  
        legend.position = "none",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", size = 0.5),
        axis.ticks.y = element_line(color = "black", size = 0.5)) + 
  coord_cartesian(clip = "off")

```

```{r}

# Panel 4: Species Observed
plot4 <- ggplot(summary_data, aes(x = StreamType, y = mean_ObsSpp, fill = StreamType)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5, width = 0.4) +
  geom_errorbar(aes(ymin = mean_ObsSpp - se_ObsSpp, ymax = mean_ObsSpp + se_ObsSpp), 
                width = 0.25, position = position_dodge(0.8), color = "black") +
  labs(title = NULL, x = NULL, y = "Species Observed") +
  scale_fill_manual(values = c("darkgrey", "white")) +
  scale_x_discrete(labels = c("Control" = "Control", "Tea" = "Tea")) +   scale_y_continuous(expand = c(0, 0), breaks = seq(0, max(summary_data$mean_ObsSpp), by = 10)) +
  theme_minimal(base_size = 14) + 
  theme(axis.text.x = element_text(size = 10),
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size = 12),
        legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", size = 0.5),
        axis.ticks.y = element_line(color = "black", size = 0.5)) + 
  coord_cartesian(clip = "off")


```


```{r}

# Combine 4 plots
grid.arrange(plot1, plot2, plot3, plot4, ncol = 2, nrow = 2)

```

## Statistical Tests with the Data

# Checking for Normality (everything is normal)

```{r}
# Checking Distribution/Normality

shapiro.test(data$Family_Rich)
shapiro.test(data$EPT_Spp)
shapiro.test(data$EstSpp)
shapiro.test(data$ObsSpp)
```

# Family Richness 

```{r}
#  Family Richness
t_test_family_rich <- t.test(Family_Rich ~ StreamType, data = data, alternative = "greater")
t_test_family_rich

# The t-test for Family Richness yielded a p-value of 0.0367, that control family richness is statistically significantly greater than in tea streams.

```

```{r}

# Mann-Whitney test for Family Richness
mann_whitney_family_rich <- wilcox.test(Family_Rich ~ StreamType, data = data, alternative = "greater")
mann_whitney_family_rich

# Since the p-value is less than 0.05 (0.04071), the result indicates that there is a statistically significant difference in Family_Richness between the two groups (Forest greater than tea).

```

# EPT Richness (no significance ttest, significant Mann-Whitney)


```{r}
t_test_ept_spp <- t.test(EPT_Spp ~ StreamType, data = data, alternative = "greater")
t_test_ept_spp

# significance (p= 0.04271)
# The t-test for ept richness yielded a p-value of 0.04271, that control ept richness is statistically significantly greater than in tea streams.
```


```{r}
# Mann-Whitney test for EPT Richness
mann_whitney_ept_spp <- wilcox.test(EPT_Spp ~ StreamType, data = data, alternative = "greater")
mann_whitney_ept_spp

#The p-value = 0.02042 suggests that there is a significant difference between the "Forest" and "Tea" groups in terms of EPT richness. Since the p-value is less than 0.05, we reject the null hypothesis and conclude that the distribution of EPT species is not the same between the two stream types.

```


# Estimated Species (no significance)

```{r}
# Estimated Species
t_test_est_spp <- t.test(EstSpp ~ StreamType, data = data, alternative = "greater")
t_test_est_spp

# The t-test for EST Species Richness yielded a p-value of 0.03897, that control family richness is statistically significantly greater than in tea streams.

```

```{r}
# Mann-Whitney test for Estimated Species Richness
mann_whitney_est_spp <- wilcox.test(EstSpp ~ StreamType, data = data, alternative = "greater")
mann_whitney_est_spp

# The mann-whitney for est species Richness yielded a p-value of 0.02857, that control est species richness is statistically significantly greater than in tea streams.

```


# Observed Species

```{r}
# Observed Species
t_test_obs_spp <- t.test(ObsSpp ~ StreamType, data = data, alternative = "greater")
t_test_obs_spp

# The t-test for observed species yielded a p-value of 0.0338, that control family richness is statistically significantly greater than in tea streams.

```

```{r}
# Mann-Whitney test for Species Observed
mann_whitney_obs_spp <- wilcox.test(ObsSpp ~ StreamType, data = data, alternative = "greater")
mann_whitney_obs_spp

# The mann-whitney for species observed yielded a p-value of 0.04071, that control species observed is statistically significantly greater than in tea streams.

```


## Abundance (no significance)

```{r}
# Abundance
t_test_abund <- t.test(Abund ~ StreamType, data = data, alternative = "greater")
t_test_abund

# The t-test for abundance yielded a p-value of 0.05452, that control family richness is marginally greater than in tea streams, not quite significant.

```


```{r}
# Mann-Whitney test for Abundance
mann_whitney_abund <- wilcox.test(Abund ~ StreamType, data = data, alternative = "greater")
mann_whitney_abund

#Since the p-value is marginally greater than 0.05, we fail to reject the null hypothesis, there is technically no difference between the means.


```

# Dissolved Oxygen (no significance ttest, significant Mann Whitney)

```{r}

t_test_do <- t.test(DO ~ StreamType, data = data, alternative = "greater")
t_test_do

# The t-test for dissolved oxygen yielded a p-value of 0.03785, that control dissolved oxygen is statistically significantly greater than in tea streams.

```


```{r}
# Mann-Whitney test for Dissolved Oxygen
mann_whitney_do <- wilcox.test(DO ~ StreamType, data = data, alternative = "greater")
mann_whitney_do

# Reject the null hypothesis: Since the p-value (0.01429) is less than 0.05, we reject the null hypothesis and conclude that there is a statistically significant difference in the distributions of DO between the "Forest" and "Tea" groups.

```

## Linear Regressions with Est. Spp (no significant relationships)

# DO

```{r}

lm_do <- lm(DO ~ EstSpp, data = data)

summary(lm_do)
```

```{r}
ggplot(data, aes(x = EstSpp, y = DO)) +
  geom_point() +  # Plot the data points
  geom_smooth(method = "lm", col = "blue") +  # Add the linear regression line
  labs(title = "Linear Regression of DO on EstSPP",
       x = "Estimated Species (EstSpp)",
       y = "Dissolved Oxygen (DO)") +
  theme_minimal()
```

# Elevation

```{r}

lm_elev <- lm(Elevation ~ EstSpp, data = data)

summary(lm_elev)
```

```{r}
ggplot(data, aes(x = EstSpp, y = Elevation)) +
  geom_point() +  # Plot the data points
  geom_smooth(method = "lm", col = "blue") +  # Add the linear regression line
  labs(title = "Linear Regression of elevation on EstSpp",
       x = "Estimated Species (EstSpp)",
       y = "Elevation") +
  theme_minimal()
```
#Stream Width

```{r}

lm_width <- lm(Stream_width ~ EstSpp, data = data)

summary(lm_width)

```


```{r}
ggplot(data, aes(x = EstSpp, y = Stream_width)) +
  geom_point() +  # Plot the data points
  geom_smooth(method = "lm", col = "blue") +  # Add the linear regression line
  labs(title = "Linear Regression of stream width on EstSpp",
       x = "Estimated Species (EstSpp)",
       y = "Stream Width") +
  theme_minimal()
```

# Canopy

```{r}

lm_canopy <- lm(Canopy ~ EstSpp, data = data)

summary(lm_canopy)
```

```{r}
ggplot(data, aes(x = EstSpp, y = Canopy)) +
  geom_point() +  # Plot the data points
  geom_smooth(method = "lm", col = "blue") +  # Add the linear regression line
  labs(title = "Linear Regression of canopy on EstSpp",
       x = "Estimated Species (EstSpp)",
       y = "canopy") +
  theme_minimal()
```

# Flow rate

```{r}

lm_flow <- lm(Flow_rate ~ EstSpp, data = data)

summary(lm_flow)

```

```{r}
ggplot(data, aes(x = EstSpp, y = Flow_rate)) +
  geom_point() +  # Plot the data points
  geom_smooth(method = "lm", col = "blue") +  # Add the linear regression line
  labs(title = "Linear Regression of flow rate on EstSpp",
       x = "Estimated Species (EstSpp)",
       y = "Flow Rate") +
  theme_minimal()
```


