---
title: "Full Anna"
format: html
editor: visual
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  echo = FALSE,      # Hide the code
  warning = FALSE,   # Suppress warnings
  message = FALSE    # Suppress messages
)

```

```{r}

library(tidyverse)
library(readr)
library(dplyr)
library(tibble)

```

```{r}

bukanna <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vQ39eNa8kt_kYh0ilq_ecR7SdEzLwjhQKZZHR1sItT1w4GzEiTJMogEkVV76Gx2e6wUnd6kVpg9sIkn/pub?gid=0&single=true&output=csv"

bukanna <- read_csv(bukanna)

bukanna$Date <- as.Date(bukanna$Date, format = "%m/%d/%Y")

```

**Multiple Regression**

```{r}

model <- lm(kd ~ Photic_TSS + Photic_CDOM + Photic_CHLa, data = bukanna)

summary(model)

```

**AIC**

```{r}


# Define the full dataset and predictors
response <- "kd"
predictors <- c("Photic_TSS", "Photic_CDOM", "Photic_CHLa")  # Adjust predictors based on your dataset

# Generate all combinations of predictors
all_combinations <- unlist(
  lapply(0:length(predictors), function(x) combn(predictors, x, simplify = FALSE)), 
  recursive = FALSE
)

# Fit models for each combination and calculate AIC
model_results <- lapply(all_combinations, function(predictor_set) {
  # Create the formula for the current combination of predictors
  formula <- as.formula(paste(response, "~", ifelse(length(predictor_set) == 0, "1", paste(predictor_set, collapse = " + "))))
  
  # Fit the linear model
  model <- lm(formula, data = bukanna)
  
  # Return the model, predictors, and AIC
  list(
    predictors = predictor_set,
    model = model,
    AIC = AIC(model)
  )
})

# Convert results to a dataframe for easier analysis
results_df <- do.call(rbind, lapply(model_results, function(x) {
  data.frame(
    Predictors = paste(x$predictors, collapse = " + "),
    AIC = x$AIC
  )
}))

# Find the model with the lowest AIC
min_AIC <- min(results_df$AIC)

# Calculate the delta AIC for each model (difference from the minimum AIC)
Annaresults_df <- results_df %>%
  mutate(Delta_AIC = AIC - min_AIC) %>%
  arrange(Delta_AIC)  # Sort by delta AIC

# Print the results with Delta AIC
print(Annaresults_df)


```

Model selection based on Akaike’s Information Criterion (AIC) indicated that Photic_CHLa alone was the best predictor of light attenuation (kd), with the lowest AIC value (–36.02). Models that included CHLa with either TSS or CDOM showed slightly higher AIC values (ΔAIC \< 2), suggesting they are also plausible but do not substantially improve model performance. Moreover, as seen in the summary table of the multiple regression, TSS and CDOM were not statistically significant (Pvalues \> 0.05) of kd across this dataset. Models without CHLa—including those with only TSS, only CDOM, or both—had significantly higher AIC values (ΔAIC \> 4), indicating weaker support. Overall, these results suggest that CHLa is the most parsimonious and informative predictor of kd in this dataset, with little to no significant gain in explanatory power from adding other variables.
