---
title: "Correlation Assignment"
format: html
editor: visual
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  warning = FALSE,   # Suppress warnings
  message = FALSE    # Suppress messages
)

```

## Question 1

Look up the library GGally; it has a function named ggpairs(). Use that function to plot the atmospheric data from the Rice Rivers Center for the samples collected from 10 February to 15 February.

```{r}

library(tidyverse)

library(ggplot2)

library(GGally)

library(lubridate)

source("get_rice_data.R")

data <- getRiceData()

```


  
```{r}

data$Date <- ymd_hms(data$Date)

atmdf <- data |>
  filter(month(Date) == 2 & day(Date) >= 10 & day(Date) <= 15) |>
  select(AirTemp, Rain_m, WindDir, BP_HG, RelHumidity, WindSpeed_kph)

ggpairs(
  data = atmdf,
  columns = 1:ncol(atmdf)) 

```
  
## Question 2

For those atmospheric data, which pair of variables has the strongest correlation? What is the 95% confidence interval on that correlation coefficient?
  
```{r}

# Calculate correlation matrix for the variables in atmdf
cor_matrix <- cor(atmdf, use = "complete.obs")

library(dplyr)
cor_df <- as.data.frame(as.table(cor_matrix)) %>%
  filter(Var1 != Var2) %>% 
  mutate(abs_correlation = abs(Freq)) %>% 
  arrange(desc(abs_correlation)) 

strongest_pair <- cor_df[1, c("Var1", "Var2")]
strongest_corr_value <- cor_df[1, "Freq"]

cor_test_result <- cor.test(atmdf[[strongest_pair$Var1]], atmdf[[strongest_pair$Var2]])

cat("The strongest correlation is between", strongest_pair$Var1, "and", strongest_pair$Var2, "\n")
cat("Correlation coefficient:", strongest_corr_value, "\n")
cat("95% confidence interval:", cor_test_result$conf.int[1], "to", cor_test_result$conf.int[2], "\n")

```

## Question 3

Using the first 40 observations in air temperature and barometric pressure from the Rice Center data set, determine if they are individually distributed as normal random variables. 

```{r}

# Subset the first 40 observations of AirTemp and BP_HG
sample_data <- data[1:40, ]


par(mfrow = c(1, 2))
qqnorm(sample_data$AirTemp, main = "Q-Q Plot of AirTemp")
qqline(sample_data$AirTemp, col = "red")

qqnorm(sample_data$BP_HG, main = "Q-Q Plot of BP_HG")
qqline(sample_data$BP_HG, col = "red")

shapiro_test_AirTemp <- shapiro.test(sample_data$AirTemp)
shapiro_test_BP_HG <- shapiro.test(sample_data$BP_HG)

cat("Shapiro-Wilk test for AirTemp:\n")
print(shapiro_test_AirTemp)

cat("\nShapiro-Wilk test for BP_HG:\n")
print(shapiro_test_BP_HG)


```

The Shapiro-Wilk test indicates that AirTemp is likely normally distributed (p = 0.2031), as we do not reject the null hypothesis of normality. However, BP_HG is not normally distributed (p = 0.0016), as the test shows a significant deviation from normality.


## Question 4

Given your findings in the last question, what kind of correlation statistic would be most appropriate for estimating the correlation between this subset of data?

Given that AirTemp appears to be normally distributed but BP_HG does not, a non-parametric correlation statistic would be most appropriate. The Spearman rank correlation could be a suitable choice, as it doesn't assume normality.

## Question 5

Look at a qqnorm() plot of the barometric pressure data you used in the previous example. Is there something that "looks" odd with these data? Explain why those data are the way they are.


In the Q-Q plot for BP_HG, there’s an unusual pattern where all the points fall on a straight vertical line, indicating that all values are nearly identical. This is unusual because, in typical data, you would expect more spread in the values

This pattern suggests that the BP_HG data might be nearly constant, possibly due to measurement limitations, a data entry error, or because it was recorded during a period with very stable atmospheric conditions, leading to almost identical barometric pressure readings. This lack of variability is likely why the Shapiro-Wilk test flagged BP_HG as non-normal.
  
## Question 6

Using a permutation approach, define the distribution of correlation values between the variables in #3, assuming that the NULL hypothesis is true. Plot these as a histogram and include the observed correlation. 

```{r}

set.seed(42)  # For reproducibility

observed_corr <- cor(sample_data$AirTemp, sample_data$BP_HG, method = "spearman")

n_permutations <- 1000
permuted_corrs <- numeric(n_permutations)

for (i in 1:n_permutations) {
  permuted_BP_HG <- sample(sample_data$BP_HG)
  permuted_corrs[i] <- cor(sample_data$AirTemp, permuted_BP_HG, method = "spearman")
}

hist(permuted_corrs, breaks = 30, main = "Permutation Distribution of Correlation (Null Hypothesis)",
     xlab = "Correlation", xlim = range(c(permuted_corrs, observed_corr)), col = "lightblue", border = "black")
abline(v = observed_corr, col = "red", lwd = 2, lty = 2)
legend("topright", legend = paste("Observed Correlation =", round(observed_corr, 3)),
       col = "red", lty = 2, lwd = 2)

```

