---
title: "Raster Data Homework"
format: html
editor: visual
---

## Question 1

Load the raster and point data in and crop to an appropriate size to display the locations of the sampling plots and make a label for each site.

```{r}

library(raster)
library(sf)
library( tidyverse )
library (ggplot2)

raster_url <- "https://github.com/DyerlabTeaching/Raster-Data/raw/main/data/alt_22.tif"

beetle_url <- "https://raw.githubusercontent.com/DyerlabTeaching/Raster-Data/main/data/AraptusDispersalBias.csv"

rast <- raster(raster_url)

```

```{r}

read_csv( beetle_url ) |>
  st_as_sf( coords=c("Longitude","Latitude"), crs=4326 ) -> beetles

beetles |> st_bbox()

```

```{r}
library(raster)
library(sf)

baja_extent <- extent( c(-115, -109, 23, 30 ) )
baja_extent

par(mar = c(2, 2, 1, 1)) #This changes the size of the plot within the window (bottom, left, top, right margins)

alt <- crop( rast, baja_extent )
plot(alt)
plot( beetles["Suitability"], 
      pch=16, 
      add=TRUE)

text(beetles$geometry, labels = beetles$Site, pos = 4, cex = 0.7, col = "blue")

```

## Question 2

Use the click() function to crop the raster and filter the sites to include only the sampling locations that are in the Cape regions (e.g., south of site labeled #75). Plot the raster of elevation and all cape region locales with sex-ratio represented as the size of the plot symbol

```{r}

#copy into the console!!

# par(mar = c(2, 2, 1, 1))
# alt <- crop(rast, baja_extent)
# 
# plot(alt)
# 
# cape_extent <- click(n = 2)

```

```{r}
# Crop the raster to the Cape region based on selected coordinates
cape_raster <- crop(alt, extent(cape_extent))

plot(cape_raster)


```

```{r}

alt |>
  rasterToPoints() |>
  as.data.frame() |> 
  transmute(Longitude=x,
            Latitude=y,
            Elevation=alt_22)  -> alt.df
head( alt.df )

alt.df |>
  ggplot()  + 
  geom_raster( aes( x = Longitude, 
                    y = Latitude, 
                    fill = Elevation) ) + 
  coord_equal() +
  theme_minimal() -> baja_elevation

baja_elevation
```

```{r}

baja_elevation + 
  scale_fill_gradient2( low = "darkolivegreen",
                        mid = "yellow",
                        high = "brown", 
                        midpoint = 1000 ) -> baja_map

baja_map + 
  geom_sf( aes(size = MFRatio ), 
           data = beetles, 
           color = "dodgerblue2",
           alpha = 0.75) 

```

## Question 3

The peninsula of Baja California is divided into the States of Baja California Norte and Baja California Sur. The border between these states is at 28° Latitude. Divide the sample locations into groups based on which state they are located in and plot the average sex ratio of the sample sites partitioned by each site.

```{r}

# Extract coordinates to create a new data frame
beetles_df <- st_as_sf(beetles)

# Create a new column for the state based on latitude
beetles_df <- beetles_df |>
  mutate(State = ifelse(st_coordinates(geometry)[,2] >= 28, 
                        "Baja California Norte", 
                        "Baja California Sur"))

# Calculate the average sex ratio by site and state
average_sex_ratio <- beetles_df |>
  group_by(Site, State) |>
  summarize(Average_MFRatio = mean(MFRatio, na.rm = TRUE))

ggplot(average_sex_ratio, aes(x = Site, y = Average_MFRatio, fill = State)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Sex Ratio by Site in Baja California",
       x = "Site",
       y = "Average Sex Ratio") +
  theme_minimal() +
  theme(
     legend.text = element_text(size = 4.5),
    legend.title = element_text(size = 4.5),  
    legend.key.size = unit(0.5, "cm"),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 6))


```

```{r}
# Extract coordinates to create a new data frame
beetles_df <- st_as_sf(beetles)

# Create a new column for the state based on latitude
beetles_df <- beetles_df |>
  mutate(State = ifelse(st_coordinates(geometry)[,2] >= 28, 
                        "Baja California Norte", 
                        "Baja California Sur"))

# Calculate the average sex ratio by site and state
average_sex_ratio <- beetles_df |>
  group_by(State) |>
  summarize(Average_MFRatio = mean(MFRatio, na.rm = TRUE))

ggplot(average_sex_ratio, aes(x = State, y = Average_MFRatio, fill = State)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Sex Ratio by State in Baja California",
       x = "State",
       y = "Average Sex Ratio") +
  theme_minimal() +
  theme(
     legend.text = element_text(size = 4.5),
    legend.title = element_text(size = 4.5),  
    legend.key.size = unit(0.5, "cm"),
    axis.text.x = element_text(size = 6))
```

```{r}

baja_extent <- extent( c(-115, -109, 23, 30 ) )
baja_extent

state_colors <- setNames(c("red", "blue"), unique(beetles_df$State))

par(mar = c(2, 2, 1, 1)) #This changes the size of the plot within the window (bottom, left, top, right margins)

alt <- crop( rast, baja_extent )
plot(alt)
plot( beetles["Suitability"], 
      pch=16, 
      add=TRUE)

text(beetles$geometry, labels = beetles$MFRatio, pos = 4, cex = 0.7, col = state_colors[beetles_df$State])

```

## Question 4

Is there a relationship between the observed sex ratio and the elevation at that site? Plot these, and in the text, perform a correlation test (cor.test()) and report the values as if you were writing them up in a results section.

```{r}

beet_df <- as.data.frame(beetles)  
coordinates <- st_coordinates(beetles$geometry) 

#add lat/long to beet_df
beet_df$Longitude <- coordinates[, 1]
beet_df$Latitude <- coordinates[, 2]

#get elevation from raster based on lat/long
elevation_values <- extract(rast, beet_df[, c("Longitude", "Latitude")])

#add elevations to beetles_df
beet_df$Elevation <- elevation_values


ggplot(beet_df, aes(x = Elevation, y = MFRatio)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Elevation", y = "MF Ratio", 
       title = "Relationship Between Elevation and MF Ratio") +
  theme_minimal()


# Perform the correlation test
cor_test_result <- cor.test(beet_df$MFRatio, beet_df$Elevation)

 r <- cor.response$estimate
  p_value <- cor.response$p.value
  
cat("r =", r, "\n")
cat("P =", p_value, "\n")

```
