---
title: "Regression Homework Activity"
author: Jackson Barnett
format: 
  html:
    code-tools: true
    toc: false
---


```{r setup, include=FALSE}
library( tidyverse )
library( nycflights13 )
```

Since regression analysis is such a fundamental component of inferences, I thought it would be a good idea to try to work through a few more examples to give you some more practice.  

For this we will use some data on flights into and out of Newark (EWR), JFK, and La Guardia airports in NYC metropolitan area.

```{r}
summary( flights ) 
```

We also have some data related to local weather conditions.

```{r}
summary( weather )
```

I'm going to make a synthetic data set consisting of flights out of JFK for the *ExpressJet* airlines.  Since the weather data is on the hour, I also make a synthetic varible `Month.Day.Hour` so that I can join the weather data to this one.

```{r}
flights %>%
  filter( origin == "JFK", carrier == "EV") %>%
  mutate( DateTime = paste(month,day,hour, sep=".")) %>%
  select( DateTime, 
          `Departure Delay` = dep_delay,
          `Departure Time` = sched_dep_time) %>%
  droplevels() -> df.jfk 
summary( df.jfk )
```

Now I'll do the same with the weather data.

```{r}
weather %>%
  filter( origin == "JFK") %>% 
  mutate( DateTime = paste(month,day,hour, sep=".")) %>%
  select( -origin, -year, -month, -day, -hour, -time_hour ) -> df.weather 
summary( df.weather )
```

Now, I'm going to `merge` these two `data.frame` objects using the common `key` I made `DateTime`.

```{r}
merge( df.jfk, df.weather, by="DateTime" ) |>
  select( -DateTime ) -> df 

summary( df )

```



## Activity

What features of this data set may impact the departure delay for flights coming out of JFK on *ExpressJet* Airlines?


## Abstract

This study investigated the factors influencing departure delays for ExpressJet flights departing from JFK airport by performing linear regression analyses. Each variable in the dataset was tested as a potential predictor of departure delay, both individually and in combination. Single regression analysis with departure time as the predictor resulted in a significant relationship
ğ‘
<
0.05
p<0.05), with an equation of 
ğ‘¦
=
0.0213
ğ‘¥
âˆ’
7.6108
y=0.0213xâˆ’7.6108 and an 
ğ‘…
2
R 
2
  of 0.034, indicating a weak explanation of variance.

Multiple regressions revealed that significant predictors of departure delay (
ğ‘
<
0.05
p<0.05) included dew point, humidity, wind direction, wind speed, precipitation, pressure, and visibility, in addition to departure time. Among these, visibility (
ğ‘…
2
=
0.0266
R 
2
 =0.0266) and humidity (
ğ‘…
2
=
0.0188
R 
2
 =0.0188) showed the strongest relationships. However, all 
ğ‘…
2
R 
2
  values were low, suggesting that while these factors contribute to delays, other unmeasured variables might play a more substantial role. Overall, the study highlights environmental and temporal variables as contributing factors to departure delays but indicates the need for further research to improve predictive models.
  
## Regressions

One Regression at a time:

```{r}
fit <- lm(`Departure Delay` ~ `Departure Time`, data = df)
summary_fit <- summary(fit)

# Extract coefficients
intercept <- round(coef(fit)[1], 4)
slope <- round(coef(fit)[2], 4)

# Extract R-squared and p-value
r_squared <- round(summary_fit$r.squared, 4)
p_value <- round(summary_fit$coefficients[2, 4], 4)

# Create the regression equation
regression_equation <- paste0(
  "Response: Departure Delay vs Departure Time\n",
  "y = ", slope, "x + ", intercept, "\n",
  "RÂ² = ", r_squared, ", p = ", p_value
)

# Print the results
cat(regression_equation, "\n\n")


```

Multiple Regressions at a time:

```{r}

predictors <- setdiff(names(df), c("Departure Delay", "Departure Time"))

for (predictor in predictors) {
  
  formula <- as.formula(paste("`Departure Delay` ~", predictor))
  
  fit <- lm(formula, data = df)
  summary_fit <- summary(fit)
  
  # Extract coefficients
  intercept <- round(coef(fit)[1], 4)
  slope <- round(coef(fit)[2], 4)
  
  # Extract R-squared and p-value
  r_squared <- round(summary_fit$r.squared, 4)
  p_value <- round(summary_fit$coefficients[2, 4], 4)
  
  # Create the regression equation
  regression_equation <- paste0(
    "Response: Departure Delay vs ", predictor, "\n",
    "y = ", slope, "x", " + ", intercept, "\n",
    "RÂ² = ", r_squared, ", p = ", p_value
  )
  
  # Print the results
  cat(regression_equation, "\n\n")
}

```


The variables with significant P-values (below 0.05) are: Departure Time, Dew Point, Humidity, Wind Direction, Wind Speed, Precipitation, Pressure, and Visibility. This indicates that all of these variables have at least some influence over Departure Delays for ExpressJet flights coming out of JFK. The R2 values vary across predictors, with the largest explanation of delay variability coming from Departure Time and Visibility.

## AIC Just for practice


```{r}

library(dplyr)

response <- "Departure Delay"
predictors <- setdiff(names(df), response)  # Exclude the response variable

# Generate all combinations of predictors
all_combinations <- unlist(
  lapply(0:length(predictors), function(x) combn(predictors, x, simplify = FALSE)), 
  recursive = FALSE
)

model_results <- lapply(all_combinations, function(predictor_set) {
  # Wrap predictor names and response in backticks
  predictor_set_escaped <- ifelse(length(predictor_set) > 0, 
                                  paste0("`", predictor_set, "`"), 
                                  "1")
  formula <- as.formula(
    paste0("`", response, "` ~ ", paste(predictor_set_escaped, collapse = " + "))
  )
  
  # Fit the model
  model <- lm(formula, data = df)
  
  list(
    predictors = predictor_set,
    model = model,
    AIC = AIC(model)
  )
})

results_df <- do.call(rbind, lapply(model_results, function(x) {
  data.frame(
    Predictors = paste(x$predictors, collapse = " + "),
    AIC = x$AIC
  )
}))

min_AIC <- min(results_df$AIC)

results_df <- results_df %>%
  mutate(Delta_AIC = AIC - min_AIC) %>%
  arrange(Delta_AIC)  # Sort by delta AIC

# Print the results with Delta AIC
head(results_df, 10)


```

