---
title: "Analysis of Variance"
author: "Jackson Barnett"
subtitle: "In Class Activity"
execute: 
  echo: true
format: 
  html:
    code-tools: true
    toc: false
---

Here is some new data to play with a bit to explore one-way and two-way analysis of variance models. These data represent weight data associated with three different diet types. There are 76 observations of individuals with the following characteristics:

-   Gender (0/1)
-   Age (earth years. 🤓)
-   Height (cm)
-   Pre.weight & weight6weeks (kg)
-   Diet (1,2,3)

The data are located as a CSV file named `DietData.csv` linked from the Canvas page. Load the data in and format things so they look correctly to you.

```{r}

library(tidyverse)
library(stringr)
library(readr)

```

```{r}

dietdata <- read_csv (  "https://docs.google.com/spreadsheets/d/e/2PACX-1vS74VGBIAxot-huGUwwHADBL7gzSDC2-AbB98O9O2_PqAhi4d2d4f5eytBi41Yw4JXIW2j7VHGehTrf/pub?gid=0&single=true&output=csv" )

```


## Exploratory Data Anlsysis

One of the first things to do is to look at the data and see if there are any obvious things. Go ahead and explore these data visually. What do you see?

It seems as though all of the data are normally distributed meaning that t tests, anovas, and any other statistical test can be applied confidently to this data. We do have to be sure that the Person, Gender, and Diet values are recognized as categories and not a continuous variable as that could cause issues when running statistical tests (this is done in later sections)

```{r}

hist(dietdata$pre.weight, breaks = 15, probability = TRUE, 
     main = "Histogram of Pre-Weight", 
     xlab = "Values", col = "lightblue", border = "black")
curve(dnorm(x, mean = mean(dietdata$pre.weight), sd = sd(dietdata$pre.weight)), 
      col = "red", lwd = 2, add = TRUE)


hist(dietdata$Age, breaks = 15, probability = TRUE, 
     main = "Histogram of Age", 
     xlab = "Values", col = "lightblue", border = "black")
curve(dnorm(x, mean = mean(dietdata$Age), sd = sd(dietdata$Age)), 
      col = "red", lwd = 2, add = TRUE)


hist(dietdata$Height, breaks = 15, probability = TRUE, 
     main = "Histogram of Height", 
     xlab = "Values", col = "lightblue", border = "black")
curve(dnorm(x, mean = mean(dietdata$Height), sd = sd(dietdata$Height)), 
      col = "red", lwd = 2, add = TRUE)


hist(dietdata$weight6weeks, breaks = 15, probability = TRUE, 
     main = "Histogram of weight6weeks", 
     xlab = "Values", col = "lightblue", border = "black")
curve(dnorm(x, mean = mean(dietdata$weight6weeks), sd = sd(dietdata$weight6weeks)), 
      col = "red", lwd = 2, add = TRUE)



```


## Estimating Mean Values

Make a table of Weight Loss by gender and diet.

```{r}

loss <- dietdata |>
  group_by(Diet) |>
  mutate(WeightLoss = pre.weight - weight6weeks) |>
  select( gender, Diet, WeightLoss)

head(loss)

hist(loss$WeightLoss, breaks = 15, probability = TRUE, 
     main = "Histogram of Weight Loss", 
     xlab = "Values", col = "lightblue", border = "black")
curve(dnorm(x, mean = mean(loss$WeightLoss), sd = sd(loss$WeightLoss)), 
      col = "red", lwd = 2, add = TRUE)

```

## 1-Way Analysis of Variance

*Diet Issues:*

The underlying linear model.

$$
y_{ij} = \mu + \tau_{Diet, i} + \epsilon_j
$$

Test the null hypothesis, $H_O:$ *There is no effect of diet on weight loss* (e.g., $\tau_{Diet-1} = \tau_{Diet-2} = \tau_{Diet-3} = 0.0$). Is there evidence for one diet producing more weight loss than the others? Create an `aov` analysis and assign it to the variable `fit.diet` and examine its contents.

```{r}
# Define mode?
loss$Diet <- as.factor(loss$Diet)

fit.diet <- aov(WeightLoss ~ Diet, data = loss)

anova(fit.diet)
```

There is strong evidence (P = 0.006596) to reject the null hypothesis. This means that at least one of the diets leads to significantly different weight loss compared to the others.

Are they all significantly different? Try the `TukeyHSD()` Interpret the results.

The Tukey test shows that diet 3 is statistically different than diets 1 and 2, but diets 1 and 2 are not different from each other. This indicates that diet 3 is the program causing such variation in the means as shown in the anova test. 

```{r}
# Posthoc test

tuk <- TukeyHSD(fit.diet)
tuk

```

How much of the variation is explained? If you notice when you do a summary from a `lm()` (regression) model, it gives you the $R^2$ values directly (remember $R^2 = \frac{SS_{model}}{SS_{Error}}$). Does `summary()` of your `aov` model give you that?

```{r}

summary(fit.diet)

```

Since I asked the question, the answer is probably no. Why does it not do this? Probably for historical reasons, which are a bit of a pain in the backside. That being said, there are some `tidy` ways to fix this issue. I'm going to use the `broom` package which allows us to clean up (or tidy if you will) the model objects. This will take the model object and pull out *all* the ANOVA table stuff and put it into a `tibble`.

```{r}
library( broom )
tidy_diet <- tidy( fit.diet )
tidy_diet
```

Now, since it is all easily accessible, we can calculate the $R^2$ from the new model output.

```{r}
# Estimate the variance explained from the raw sums of squares
r2_Diet <- tidy_diet$sumsq[1] / sum( tidy_diet$sumsq )
r2_Diet
```

*Gender:*

The underlying linear model.

$y_{ij} = \mu + \tau_{gender, i} + \epsilon_j$

Independent of the diet, test the null hypothesis $H_O:$ *There is no difference in weight loss between genders* (e.g., \$\tau\*{gender-0} =\* \tau{gender-2} = 0.0 \$). Is there evidence for one gender being significantly different than another? How much of the variation is explained (another $R^2$ by gender)?

```{r}

loss$gender <- as.factor(loss$gender)

fit.gender <- aov(WeightLoss ~ gender, data = loss)

anova(fit.gender)

```

How do you interpret these results thus far?

It does not seem as though gender has a significant effect on weight loss (P = 0.8348)

# Do genders respond differently to diets?

$y_{ijk} = \mu + \tau_{Diet,i} + \tau_{gender,j} + \epsilon_k$

So here we do something a bit different. We want to simultaneously ask the following questions:

-   Do diets influence weight loss?
-   Do genders influence weight loss?
-   Is there an interaction where different genders respond differently to different diets?

In $R$, this is done as:

```{r}

tidy_diet <- tidy( fit.diet )
tidy_diet

r2_Diet <- tidy_diet$sumsq[1] / sum( tidy_diet$sumsq )
r2_Diet

```

```{r}
# gender model

tidy_gend <- tidy( fit.gender )
tidy_gend

r2_gend <- tidy_gend$sumsq[1] / sum( tidy_gend$sumsq )
r2_gend

```

```{r}
# interaction model

fit.int <- aov(WeightLoss ~ Diet * gender, data = loss)

anova_table <- anova(fit.int)

tidy_int <- tidy( fit.int )
tidy_int

```

What is the $R^2$ for this model?

```{r}
#based on after class conversation

ss_interaction <- anova_table["Diet:gender", "Sum Sq"]
ss_total <- sum(anova_table[["Sum Sq"]])

r2_int <- ss_interaction / ss_total

r2_int


```


# Which Model is Best?

How would you compare the models you generated? How do you interpret the findings?

*Explain, in words, your findings and interpretation of these findings*


The diet model was significant and able to explain approximately 12.85% of the variation in the weight loss data while the gender model was insignificant and could not explain much additional variation regardless. The model taking into account the influence of gender and diet interacting, however, was significant and explained approximately 7.2% of the variation in weight loss data. This indicates that, while gender alone is not influential for the amount of weight lost, there are significant interactions between diet types and genders that can affect weight loss. 


