---
title: "Raster Data Homework"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)

```

## Question 1

Load the raster and point data in and crop to an appropriate size to display the locations of the sampling plots and make a label for each site.

```{r}

library(raster)
library(sf)
library( tidyverse )
library (ggplot2)

raster_url <- "https://github.com/DyerlabTeaching/Raster-Data/raw/main/data/alt_22.tif"

beetle_url <- "https://raw.githubusercontent.com/DyerlabTeaching/Raster-Data/main/data/AraptusDispersalBias.csv"

rast <- raster(raster_url)

```

```{r}

read_csv( beetle_url ) |>
  st_as_sf( coords=c("Longitude","Latitude"), crs=4326 ) -> beetles

beetles |> st_bbox()

```

```{r}
library(raster)
library(sf)

baja_extent <- extent( c(-115, -109, 23, 30 ) )
baja_extent


par(mar = c(2, 2, 1, 1)) #This changes the size of the plot within the window (bottom, left, top, right margins)

alt <- crop( rast, baja_extent )
plot(alt)
plot( beetles["Suitability"], 
      pch=16, 
      add=TRUE)

text(beetles$geometry, labels = beetles$Site, pos = 4, cex = 0.7, col = "blue")

```

## Question 2

Use the click() function to crop the raster and filter the sites to include only the sampling locations that are in the Cape regions (e.g., south of site labeled #75). Plot the raster of elevation and all cape region locales with sex-ratio represented as the size of the plot symbol

Normally, we would copy this into the console to define a new extent using the click function:

alt \<- crop( rast, baja_extent ) plot(alt)

text(beetles$geometry, labels = beetles$Site, pos = 4, cex = 0.7, col = "blue")

cape_extent \<- click(n = 2)

The code below is a rendered approximation of the output that would be seen when the click function is used in the console. The cape_extent would be used for this typically.

```{r}
clickbeet <-  st_as_sf(beetles)
  
newclickbeet <- clickbeet[st_coordinates(clickbeet)[, 2] < 24.6, ]

cape_extent <- extent( c(-115, -109, 23, 24.6 ))

alt <- crop( rast, cape_extent )
plot(alt)
plot(newclickbeet["Suitability"], pch = 16, add = TRUE)

text(st_geometry(newclickbeet), 
     labels = newclickbeet$Site, 
     pos = 4, 
     cex = 0.7, 
     col = "blue")

```

```{r}

alt |>
  rasterToPoints() |>
  as.data.frame() |> 
  transmute(Longitude=x,
            Latitude=y,
            Elevation=alt_22)  -> alt.df
head( alt.df )

alt.df |>
  ggplot()  + 
  geom_raster( aes( x = Longitude, 
                    y = Latitude, 
                    fill = Elevation) ) + 
  coord_equal() +
  theme_minimal() -> baja_elevation

baja_elevation
```

```{r}

baja_elevation + 
  scale_fill_gradient2( low = "darkolivegreen",
                        mid = "yellow",
                        high = "brown", 
                        midpoint = 1000 ) -> baja_map

baja_map + 
  geom_sf( aes(size = MFRatio ), 
           data = newclickbeet, 
           color = "dodgerblue2",
           alpha = 0.75) 

```

## Question 3

The peninsula of Baja California is divided into the States of Baja California Norte and Baja California Sur. The border between these states is at 28° Latitude. Divide the sample locations into groups based on which state they are located in and plot the average sex ratio of the sample sites partitioned by each site.

This question seemed like it could have been interpereted in a few different ways, so there are three different plots of the average sex ratio partitioned by the sample sites below. I am assuming the third plot is the one that was expected.

```{r}
# Extract coordinates to create a new data frame
beetles_df <- st_as_sf(beetles)

# Create a new column for the state based on latitude
beetles_df <- beetles_df |>
  mutate(State = ifelse(st_coordinates(geometry)[,2] >= 28, 
                        "Baja California Norte", 
                        "Baja California Sur"))

# Calculate the average sex ratio by site and state
average_sex_ratio <- beetles_df |>
  group_by(State) |>
  summarize(Average_MFRatio = mean(MFRatio, na.rm = TRUE))

ggplot(average_sex_ratio, aes(x = State, y = Average_MFRatio, fill = State)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Sex Ratio by State in Baja California",
       x = "State",
       y = "Average Sex Ratio") +
  theme_minimal() +
  theme(
     legend.text = element_text(size = 4.5),
    legend.title = element_text(size = 4.5),  
    legend.key.size = unit(0.5, "cm"),
    axis.text.x = element_text(size = 6))
```

```{r}

# Calculate the average sex ratio by site and state
average_sex_ratio <- beetles_df |>
  group_by(Site, State) |>
  summarize(Average_MFRatio = mean(MFRatio, na.rm = TRUE))

ggplot(average_sex_ratio, aes(x = Site, y = Average_MFRatio, fill = State)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Sex Ratio by Site in Baja California",
       x = "Site",
       y = "Average Sex Ratio") +
  theme_minimal() +
  theme(
     legend.text = element_text(size = 4.5),
    legend.title = element_text(size = 4.5),  
    legend.key.size = unit(0.5, "cm"),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 6))


```

```{r}
baja_extent <- extent( c(-115, -109, 23, 30 ) )
baja_extent

state_colors <- setNames(c("red", "blue"), unique(beetles_df$State))

par(mar = c(2, 2, 1, 1)) #This changes the size of the plot within the window (bottom, left, top, right margins)

alt <- crop( rast, baja_extent )
plot(alt)
plot( beetles["Suitability"],
      col = "black",
      pch=16, 
      add=TRUE)

text(beetles$geometry, labels = beetles$MFRatio, pos = 4, cex = 0.7, col = state_colors[beetles_df$State])

legend("bottomleft",                       
       legend = names(state_colors),     
       col = state_colors,               
       pch = 16,                         
       title = "State") 

```

## Question 4

Is there a relationship between the observed sex ratio and the elevation at that site? Plot these, and in the text, perform a correlation test (cor.test()) and report the values as if you were writing them up in a results section.

```{r}

# beetles <- st_transform(beetles, crs(rast))
# 
# beet_df <- as.data.frame(beetles)  
# coordinates <- st_coordinates(beetles$geometry)
# 
# #add lat/long to beet_df
# beet_df$Longitude <- coordinates[, 1]
# beet_df$Latitude <- coordinates[, 2]
# 
# 
# #get elevation from raster based on lat/long
# elevation_values <- extract(rast, as.matrix(beet_df[, c("Longitude", "Latitude")]))
# 
# 
# #add elevations to beetles_df
# beet_df$Elevation <- elevation_values
# 
# 
# ggplot(beet_df, aes(x = Elevation, y = MFRatio)) +
#   geom_point() +
#   geom_smooth(method = "lm", se = FALSE, color = "blue") +
#   labs(x = "Elevation", y = "MF Ratio", 
#        title = "Relationship Between Elevation and MF Ratio") +
#   theme_minimal()
# 
# 
# # Perform the correlation test
# cor_test_result <- cor.test(beet_df$MFRatio, beet_df$Elevation)
# 
#  r <- cor.response$estimate
#   p_value <- cor.response$p.value
#   
# cat("r =", r, "\n")
# cat("P =", p_value, "\n")

```

```{r}

library(raster)
library(sp)

beetles <- st_transform(beetles, crs(rast))

beet_df <- as.data.frame(beetles)  
coordinates <- st_coordinates(beetles$geometry)

#add lat/long to beet_df
beet_df$Longitude <- coordinates[, 1]
beet_df$Latitude <- coordinates[, 2]

# Convert coordinates to cell indices
cell_indices <- cellFromXY(rast, beet_df[, c("Longitude", "Latitude")])

# Get elevation values directly using the cell indices
elevation_values <- rast[cell_indices]

# Add elevations to beet_df
beet_df$Elevation <- elevation_values

# Continue with ggplot2 plotting
ggplot(beet_df, aes(x = Elevation, y = MFRatio)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Elevation", y = "MF Ratio", 
       title = "Relationship Between Elevation and MF Ratio") +
  theme_minimal()

# Perform the correlation test
cor_test_result <- cor.test(beet_df$MFRatio, beet_df$Elevation)

# Access correlation results correctly
r <- cor_test_result$estimate
p_value <- cor_test_result$p.value
  
cat("r =", r, "\n")
cat("P =", p_value, "\n")
```

## Results

The correlation analysis between elevation and the male-to-female ratio (MFRatio) yielded a correlation coefficient of r=−0.103, indicating a very weak negative correlation between the two variables. This suggests that as elevation increases, there is a slight tendency for the MFRatio to decrease; however, the relationship is negligible and likely not meaningful in practical terms.

Further examination of the statistical significance of this correlation revealed a p-value of P=0.580. This value exceeds the conventional threshold of 0.05, leading us to fail to reject the null hypothesis. Consequently, we conclude that there is insufficient evidence to support a statistically significant relationship between elevation and MFRatio in the sampled data. In summary, variations in elevation do not appear to have a substantial effect on the MFRatio within this dataset.
