---
title: "Longitudinal Analysis"
format: html
editor: visual
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  echo = FALSE,      # Hide the code
  warning = FALSE,   # Suppress warnings
  message = FALSE    # Suppress messages
)

```


```{r}

library(tidyverse)

url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vQC8V8v0yeGHO-Ojzlf2jPfzfEoXLJPmX2ii_U2-kjh3ulpj1zAv1mhja2CYrKDm0UoT2TLlPiwqeiU/pub?gid=1048641840&single=true&output=csv"

long <- read_csv(url)

```

## Mean CHLa Regressions

```{r}

library(ggplot2)
library(dplyr)

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  summarize(
    intercept = coef(lm(CHLa_mean ~ `Distance (m)`))[1],
    slope = coef(lm(CHLa_mean ~ `Distance (m)`))[2],
    r_squared = summary(lm(CHLa_mean ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(CHLa_mean ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4),"x ", "+ ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = CHLa_mean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "CHLa_mean", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  ) 

```

## Median CHLa Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  summarize(
    intercept = coef(lm(CHLa_median ~ `Distance (m)`))[1],
    slope = coef(lm(CHLa_median ~ `Distance (m)`))[2],
    r_squared = summary(lm(CHLa_median ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(CHLa_median ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = CHLa_median)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "CHLa_median", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )



```

## CHLa Slope Regression

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>% 
  filter(!all(is.na(CHLa_slope)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(CHLa_slope ~ `Distance (m)`))[1],
    slope = coef(lm(CHLa_slope ~ `Distance (m)`))[2],
    r_squared = summary(lm(CHLa_slope ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(CHLa_slope ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 6), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = CHLa_slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "CHLa_slope", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## Mean DIN Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(DIN_mean)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(DIN_mean ~ `Distance (m)`))[1],
    slope = coef(lm(DIN_mean ~ `Distance (m)`))[2],
    r_squared = summary(lm(DIN_mean ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(DIN_mean ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = DIN_mean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "DIN_mean", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## Median DIN Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(DIN_median)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(DIN_median ~ `Distance (m)`))[1],
    slope = coef(lm(DIN_median ~ `Distance (m)`))[2],
    r_squared = summary(lm(DIN_median ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(DIN_median ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = DIN_median)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "DIN_median", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## DIN Slope Regression

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>% 
  filter(!all(is.na(DIN_slope)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(DIN_slope ~ `Distance (m)`))[1],
    slope = coef(lm(DIN_slope ~ `Distance (m)`))[2],
    r_squared = summary(lm(DIN_slope ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(DIN_slope ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 6), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = DIN_slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "DIN_slope", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## Mean Total Nitrogen Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(N_mean)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(N_mean ~ `Distance (m)`))[1],
    slope = coef(lm(N_mean ~ `Distance (m)`))[2],
    r_squared = summary(lm(N_mean ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(N_mean ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = N_mean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "N_mean", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## Median Total Nitrogen Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(N_median)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(N_median ~ `Distance (m)`))[1],
    slope = coef(lm(N_median ~ `Distance (m)`))[2],
    r_squared = summary(lm(N_median ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(N_median ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = N_median)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "N_median", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```


## Total Nitrogen Slope Regression

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>% 
  filter(!all(is.na(N_slope)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(N_slope ~ `Distance (m)`))[1],
    slope = coef(lm(N_slope ~ `Distance (m)`))[2],
    r_squared = summary(lm(N_slope ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(N_slope ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 6), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = N_slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "N_slope", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```


## Mean Total Phosphorous Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(P_mean)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(P_mean ~ `Distance (m)`))[1],
    slope = coef(lm(P_mean ~ `Distance (m)`))[2],
    r_squared = summary(lm(P_mean ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(P_mean ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = P_mean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "P_mean", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```


## Median Total Phosphorous Regression


```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(P_median)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(P_median ~ `Distance (m)`))[1],
    slope = coef(lm(P_median ~ `Distance (m)`))[2],
    r_squared = summary(lm(P_median ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(P_median ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = P_median)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "P_median", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## Total Phosphorous Slope Regression

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>% 
  filter(!all(is.na(P_slope)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(P_slope ~ `Distance (m)`))[1],
    slope = coef(lm(P_slope ~ `Distance (m)`))[2],
    r_squared = summary(lm(P_slope ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(P_slope ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 6), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = P_slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "P_slope", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## Mean Sechhi Depth Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(Secchi_mean)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(Secchi_mean ~ `Distance (m)`))[1],
    slope = coef(lm(Secchi_mean ~ `Distance (m)`))[2],
    r_squared = summary(lm(Secchi_mean ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(Secchi_mean ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = Secchi_mean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "Secchi_mean", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```


## Median Secchi Depth Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(Secchi_median)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(Secchi_median ~ `Distance (m)`))[1],
    slope = coef(lm(Secchi_median ~ `Distance (m)`))[2],
    r_squared = summary(lm(Secchi_median ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(Secchi_median ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = Secchi_median)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "Secchi_median", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```


## Secchi Slope Regression

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>% 
  filter(!all(is.na(Secchi_slope)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(Secchi_slope ~ `Distance (m)`))[1],
    slope = coef(lm(Secchi_slope ~ `Distance (m)`))[2],
    r_squared = summary(lm(Secchi_slope ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(Secchi_slope ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 6), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = Secchi_slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "Secchi_slope", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )



```

## Mean TSS Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(TSS_mean)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(TSS_mean ~ `Distance (m)`))[1],
    slope = coef(lm(TSS_mean ~ `Distance (m)`))[2],
    r_squared = summary(lm(TSS_mean ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(TSS_mean ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = TSS_mean)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "TSS_mean", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## Median TSS Regressions

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>%
  filter(!all(is.na(TSS_median)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(TSS_median ~ `Distance (m)`))[1],
    slope = coef(lm(TSS_median ~ `Distance (m)`))[2],
    r_squared = summary(lm(TSS_median ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(TSS_median ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 4), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = TSS_median)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "TSS_median", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

## TSS Slope Regression

```{r}

model_data <- long %>%
  group_by(`Lake/Reservoir`) %>% 
  filter(!all(is.na(TSS_slope)) & !all(is.na(`Distance (m)`))) %>%  # Ensure there's data to model
  summarize(
    intercept = coef(lm(TSS_slope ~ `Distance (m)`))[1],
    slope = coef(lm(TSS_slope ~ `Distance (m)`))[2],
    r_squared = summary(lm(TSS_slope ~ `Distance (m)`))$r.squared,
    p_value = summary(lm(TSS_slope ~ `Distance (m)`))$coefficients[2, 4]  # p-value for the slope
  )

# Create a label with the equation, R², and p-value for each Lake/Reservoir
model_data <- model_data %>%
  mutate(label = paste0("y = ", round(slope, 6), "x + ", round(intercept, 2), "\n",
                        "R² = ", round(r_squared, 2), "\n",
                        "p = ", round(p_value, 4)))

# Plot with annotations for each Lake/Reservoir
ggplot(long, aes(x = `Distance (m)`, y = TSS_slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Line of best fit
  facet_wrap(~ `Lake/Reservoir`, scales = "free") +         # Separate plot for each lake/reservoir
  labs(x = "Distance (m)", y = "TSS_slope", title = "Scatterplots with Best Fit Lines for Each Lake/Reservoir") +
  theme_minimal() +
  geom_text(
    data = model_data,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.5,
    inherit.aes = FALSE,
    size = 3,
    color = "black"
  )


```

